Response to reviewers

### 2025-01-02 Manuscript/ecography_reponse_letter_2.tex
Update ecography_reponse_letter_2.tex

\textcolor{blue}{Lines 128-135: The capacity of the algorithm to simulate the stressor strongly depends on the distribution of the data and the assumption of independence in the simulated dimensions (as should be the case in the simulation example). Given this, it is unclear why a single trait is simulated as multidimensional. This approach feels unnecessarily complex, potentially confounding the reader and introducing covariation that should be tested. To clarify, a trait is typically defined as a single measurable characteristic of an organism, particularly when considering the definition referenced in Violle et al. (2007). The concept of multidimensionality in trait spaces usually applies when multiple traits are combined into a matrix, with each column representing a distinct trait or a principal component summarizing trait variation. However, it is not common practice to describe a single trait using multiple variables (i.e., columns or dimensions). Regarding the statement in Lines 132–135:
“These aspects are often expressed in one dimension but can in fact be described in any number of dimensions. For example, for trait defined as “leaf insertion angle”, this can measured in one dimension (an angle in degrees) or three dimensions (the same angle expressed as the trigonometric relation of three sides of a triangle).”
While this explanation clarifies the authors’ definition of a “multidimensional trait,” it appears to complicate the methodology. If traits are often expressed in a single dimension (e.g., angle in degrees), I do not see the necessity to simulate them as “a 2, 4 or 8 dimensional trait, in other words a Brownian Motion simulated in a 2, 4 or 8 dimensions.” (lines 128-129 If I understood the text correctly, this approach results in a trait space built around a single trait defined by multiple dimensions. Instead, it would be more intuitive and broadly applicable to simulate 2, 4, or 8 separate traits. These traits, together, would define a 2-, 4-, or 8-dimensional trait space. Since these traits would be uncorrelated, they would correspond directly to the same number of principal components, aligning with the definition of dimensions provided in Supplementary Table 1.
Moreover, simulating a single multidimensional trait may inadvertently introduce covariation across dimensions. For example, using “leaf insertion angle” as an illustration, the trigonometric relationships of the triangle’s sides and angles inherently interconnect mathematically (e.g., through sine, cosine, and tangent). This interconnection could introduce correlation in the variance, even under Brownian assumptions. While this correlation may not be inherently incorrect or distinct from real ecological data, the assumption of independence in the simulation should be explicitly tested. Correlated dimensions might create clustering in the trait space, potentially biasing the application of stressors in the simulation context. 
Finally, simulating independent traits using Brownian motion without fixing the variance could also address the issue of homogeneity of variance across dimensions (Lines 463–464). Homogeneity of variance may introduce similar bias in the simulated mechanisms and it is far from approximating a set of independent trait or dimensions. I belive that simulating independent trait may mitigate that concern.}

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

\textcolor{blue}{Lines 128-135: The capacity of the algorithm to simulate the stressor strongly depends on the distribution of the data and the assumption of independence in the simulated dimensions (as should be the case in the simulation example).
Given this, it is unclear why a single trait is simulated as multidimensional.
This approach feels unnecessarily complex, potentially confounding the reader and introducing covariation that should be tested.
To clarify, a trait is typically defined as a single measurable characteristic of an organism, particularly when considering the definition referenced in Violle et al. (2007).
The concept of multidimensionality in trait spaces usually applies when multiple traits are combined into a matrix, with each column representing a distinct trait or a principal component summarizing trait variation.
However, it is not common practice to describe a single trait using multiple variables (i.e., columns or dimensions).

Regarding the statement in Lines 132–135:
\textit{“These aspects are often expressed in one dimension but can in fact be described in any number of dimensions. For example, for trait defined as “leaf insertion angle”, this can measured in one dimension (an angle in degrees) or three dimensions (the same angle expressed as the trigonometric relation of three sides of a triangle).”}
While this explanation clarifies the authors’ definition of a “multidimensional trait,” it appears to complicate the methodology.
If traits are often expressed in a single dimension (e.g., angle in degrees), I do not see the necessity to simulate them as “a 2, 4 or 8 dimensional trait, in other words a Brownian Motion simulated in a 2, 4 or 8 dimensions.” (lines 128-129 If I understood the text correctly, this approach results in a trait space built around a single trait defined by multiple dimensions.

Instead, it would be more intuitive and broadly applicable to simulate 2, 4, or 8 separate traits.
These traits, together, would define a 2-, 4-, or 8-dimensional trait space.
Since these traits would be uncorrelated, they would correspond directly to the same number of principal components, aligning with the definition of dimensions provided in Supplementary Table 1.
Moreover, simulating a single multidimensional trait may inadvertently introduce covariation across dimensions.
For example, using “leaf insertion angle” as an illustration, the trigonometric relationships of the triangle’s sides and angles inherently interconnect mathematically (e.g., through sine, cosine, and tangent).
This interconnection could introduce correlation in the variance, even under Brownian assumptions.
While this correlation may not be inherently incorrect or distinct from real ecological data, the assumption of independence in the simulation should be explicitly tested.
Correlated dimensions might create clustering in the trait space, potentially biasing the application of stressors in the simulation context. 
Finally, simulating independent traits using Brownian motion without fixing the variance could also address the issue of homogeneity of variance across dimensions (Lines 463–464).
Homogeneity of variance may introduce similar bias in the simulated mechanisms and it is far from approximating a set of independent trait or dimensions.
I belive that simulating independent trait may mitigate that concern.}

% TODO:  Change the trait definition from single to multiple. And say it's because in treats "traits" are treated as processes in n-dimensions but that we agree that this is adds extra complication here.

@@

% TODO: double check the definition and make sure it aligns with the correct one. Say we fucked up explanation because of the negative effect (i.e. removing X is the same as keeping not X).

@@

% TODO: re double checking how the scaling works and what it does to the data!

@@

% TODO: add colour legend to the plot
% TODO: distinguish between the two scalings in the figure
% TODO: expand scaling explanation in the caption

@@

% TODO: clarify the artificial extinctions in the birds dataset (yes, we always start from 118 points).
% TODO: Remove TPD metric altogether when it's not calculable?
% TODO: double check if that's what the reviewer means.

@@

% TODO: say we're keeping Box 2 because.
% TODO: but also develop points they want.

@@

% TODO: add the table to the main text
% TODO: say we keep our definition of a trait as "measurable characteristic" (but that we use the more common defintion for the simulations)

@@

% TODO: redo the supplementary plots but with different scale bars for every plot (and don't do the scaling)

@@

% TODO: double check what's in the tables
% TODO: in caption, explain how SES is calculated (reference the R package)
